---
title: "1-30.qmd"
format: html
editor: visual
---

## Notes

data.table\<- variant on a table frame, package really designed for fast operations on large data structures

tibble\<-slower than data.table all tabular data structures

\## Running Code

```{r}
library(tidyverse)
f<-"https://raw.githubusercontent.com/difiore/ada-datasets/refs/heads/main/KamilarAndCooperData.csv"
d <- read_csv(f, col_names = TRUE)
names(d)
col(d)
head(d)
rownames(d)
```

```{r}
#| echo: false
2 * 2
```

```{r}

mean(d$Brain_Size_Species_Mean, na.rm = TRUE)

attach(d)
mean(Brain_Size_Species_Mean, na.rm = TRUE)

```

attach(d) allows you to not have to do d\$Brain_Size_Species_Mean don't have too constantly specify your variables na.rm needed because mean function can't handle NA

```{r}
detach(d)
with(d,Body_mass_female_mean, na.rm = TRUE)
```

# Tidy Data

tidy data is a way to organize tabular data in a consistent data structure across packages. a table is tidy if : each table is in a column

### Not tidy Data

sometimes you will need to switch between the two between (tidy/not tidy)

#### long data

one way data cannot be tidy

# Variable summaries

```{r}
summary(d)
library(skmir)
skim(d)
```

# Box Plot

```{r}

boxplot(log(d$Body_mass_female_mean))
stripchart(log(d$Body_mass_female_mean),
           method="jitter",
           col = "blue",
           vertical = TRUE, 
           add = TRUE)
```

## Using {ggplot2}

```{r}
library(ggplot2)
p <- ggplot(data = d, aes(x="", y = log(d$Body_mass_female_mean)))
p <- p + geom_point(na.rm = TRUE)
p <- p + geom_jitter( color = "blue", width = 0.5)
p
```
